---
title       : TriangleDraw
subtitle    : An application that draws a triangle!
author      : MD
job         : CEO of TriangleDraw
logo        : triangleTitle.png
framework   : io2012        # {io2012, html5slides, shower, dzslides, landslide}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
knit        : slidify::knit2slides
---



## Basic idea

The user provides three lengths, which will be the lengths of edges comprising the triangle.



So, for example, if we choose:

```{r}
a <- 3
b <- 4
c <- 5
```

the application should return something like the following (**the following (as all figures in my presentation) is a result of an R expression which is evaluated - please see the attatched code on github**):

```{r,echo=FALSE,fig.height=3.5,fig.width=3.5}

area <- FALSE

areaTextPointXcoordinate <- function(xb,xc) {
  out <- xb
  if(xc<0) {
    out <- xb+xc
  } else if(xc>xb) {
    out <- xc
  }
  out/2
}

if( c>a+b || a>b+c || b>a+c) {
  area <- 0
  plot(c(-10,10),c(-10,10),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f) DOES NOT EXIST",a,b,c),asp=1)
  text(0,0,"Note that the condition\nc<a+b && a<b+c && b<a+c\nneeds to be satisfied")
} else {
  
  cosAlpha <-(a*a - c*c - b*b)/c/b/2
  if(a>b) {
    cosAlpha = -abs(cosAlpha)
  } else {
    cosAlpha = abs(cosAlpha)
  }
  sinAlpha <- sqrt(1 - cosAlpha*cosAlpha)
  
  xa <- 0
  ya <- 0
  xb <- c
  yb <- 0
  xc <- b*cosAlpha
  yc <- b*sinAlpha
  
  epsilon <- 0.05*min(c(a,b,c))
  plot(c(min(xc,0)-2*epsilon,max(xc,xb)+2*epsilon),c(-1,yc+2*epsilon),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f)",a,b,c),asp=1)
  polygon(x=c(0,xb,xc,0),y=c(0,yb,yc,0),col="orange",lty=3,lwd=5,border="red")
  
  text(-epsilon,0,"A")
  text(xb+epsilon,0,"B")
  text(xc,yc+epsilon,"C")
  text(xb+(xc-xb)/2+epsilon,yc/2,"a")
  text(xc/2-epsilon,yc/2,"b")
  text(xb/2,-epsilon,"c")
  if(area) {
    text(areaTextPointXcoordinate(xb,xc),yc/5,sprintf("Area = %.2f",xb*yc/2),font=4)
  }
}
```

---

## How it works

The application starts out by setting the $A$ vertex in the point $(0,0)$.
Next, the $B$ points coordinates are $(b,0)$.
To calculate the position of the $C$ point, I use the "Law of cosines": $$a^2 = c^2 + b^2 - 2bc \cos\alpha,$$ where $\alpha$ is the angle $\measuredangle CAB$. Thus, the coordinates of the $C$ points are $(b\cos\alpha,b\sin\alpha)$.

To draw the triangle I first call the ```plot()``` function to set up the canvas, and then ```polygon()``` to draw the triangle itself.

Then, the application adds text to the triangle (names of edges and vertices).

---

## How it works 2

But what if the edges do not comprise a triangle? For example: ```a <- 2; b <- 6; c <- 3```.

Then, what you should get is the following plot:

```{r,echo=FALSE,fig.height=5,fig.width=5}
a <- 2
b <- 6
c <- 3
area <- FALSE

areaTextPointXcoordinate <- function(xb,xc) {
  out <- xb
  if(xc<0) {
    out <- xb+xc
  } else if(xc>xb) {
    out <- xc
  }
  out/2
}

if( c>a+b || a>b+c || b>a+c) {
  area <- 0
  plot(c(-10,10),c(-10,10),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f) DOES NOT EXIST",a,b,c),asp=1)
  text(0,0,"Note that the condition\nc<a+b && a<b+c && b<a+c\nneeds to be satisfied")
} else {
  
  cosAlpha <-(a*a - c*c - b*b)/c/b/2
  if(a>b) {
    cosAlpha = -abs(cosAlpha)
  } else {
    cosAlpha = abs(cosAlpha)
  }
  sinAlpha <- sqrt(1 - cosAlpha*cosAlpha)
  
  xa <- 0
  ya <- 0
  xb <- c
  yb <- 0
  xc <- b*cosAlpha
  yc <- b*sinAlpha
  
  epsilon <- 0.05*min(c(a,b,c))
  plot(c(min(xc,0)-2*epsilon,max(xc,xb)+2*epsilon),c(-1,yc+2*epsilon),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f)",a,b,c),asp=1)
  polygon(x=c(0,xb,xc,0),y=c(0,yb,yc,0),col="orange",lty=3,lwd=5,border="red")
  
  text(-epsilon,0,"A")
  text(xb+epsilon,0,"B")
  text(xc,yc+epsilon,"C")
  text(xb+(xc-xb)/2+epsilon,yc/2,"a")
  text(xc/2-epsilon,yc/2,"b")
  text(xb/2,-epsilon,"c")
  if(area) {
    text(areaTextPointXcoordinate(xb,xc),yc/5,sprintf("Area = %.2f",xb*yc/2),font=4)
  }
}
```

And rightly so.

---

## Area

TriangleDraw also calculates the area of the triangle - just make sure you click the "Calculate area?" checkbox on the page. The area is then reported in the plot:

```{r,echo=FALSE,fig.height=6,fig.width=6}
a <- 3
b <- 4
c <- 5
area <- TRUE

areaTextPointXcoordinate <- function(xb,xc) {
  out <- xb
  if(xc<0) {
    out <- xb+xc
  } else if(xc>xb) {
    out <- xc
  }
  out/2
}

if( c>a+b || a>b+c || b>a+c) {
  area <- 0
  plot(c(-10,10),c(-10,10),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f) DOES NOT EXIST",a,b,c),asp=1)
  text(0,0,"Note that the condition\nc<a+b && a<b+c && b<a+c\nneeds to be satisfied")
} else {
  
  cosAlpha <-(a*a - c*c - b*b)/c/b/2
  if(a>b) {
    cosAlpha = -abs(cosAlpha)
  } else {
    cosAlpha = abs(cosAlpha)
  }
  sinAlpha <- sqrt(1 - cosAlpha*cosAlpha)
  
  xa <- 0
  ya <- 0
  xb <- c
  yb <- 0
  xc <- b*cosAlpha
  yc <- b*sinAlpha
  
  epsilon <- 0.05*min(c(a,b,c))
  plot(c(min(xc,0)-2*epsilon,max(xc,xb)+2*epsilon),c(-1,yc+2*epsilon),type="n",xlab="X",ylab="Y",main=sprintf("A triangle with edges\n(a,b,c) = (%.2f, %.2f, %.2f)",a,b,c),asp=1)
  polygon(x=c(0,xb,xc,0),y=c(0,yb,yc,0),col="orange",lty=3,lwd=5,border="red")
  
  text(-epsilon,0,"A")
  text(xb+epsilon,0,"B")
  text(xc,yc+epsilon,"C")
  text(xb+(xc-xb)/2+epsilon,yc/2,"a")
  text(xc/2-epsilon,yc/2,"b")
  text(xb/2,-epsilon,"c")
  if(area) {
    text(areaTextPointXcoordinate(xb,xc),yc/5,sprintf("Area = %.2f",xb*yc/2),font=4)
  }
}
```